Structured Versus Decorated Cospans

### 2021-06-23 cospans_new.tex
Fixed stuff at end of Section 4, after proof of Thm. 4.1.  Check it out!
tried to fix all the mistakes Kenny caught in his email of today, e.g. lots of x's

In \cref{DecCospansDoubleCat}, we construct the symmetric monoidal double category $F\lCsp$ and we define maps between decorated cospan double categories. In \cref{EquivDoubleCats}, we briefly review the structured cospans framework and prove that the double categorical versions of decorated cospans and structured cospans are isomorphic under suitable conditions. {\chris Section 3 is missing!} In \cref{spinoffs}, we establish the isomorphism between structured and decorated cospans at the level of bicategories and categories (via decategorification).  Finally, in \cref{Applications}, we describe applications, and show that open dynamical system can be described using decorated cospans but not structured cospans.

@@

We thank Daniel Cicala, Brendan Fong and Joe Moeller for helpful conversations.  We especially thank Michael Shulman for suggesting a better proof strategy.  The third author would like to thank the General Secretariat for Research  and Technology (GSRT) and the Hellenic Foundation for Research and Innovation (HFRI).

@@

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $s \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(s) \to s'$ in $F(s')$.
\end{itemize}
\end{thm}

@@

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ via \cref{lem:MonGroth}.  To obtain the desired double category we apply \cref{lem:shulman} to the functor $\Phi \maps \A/U \to \A$, which we now describe in more detail.  In the comma category $\A/U$:
\begin{itemize}
\item an object is a pair $(a \xrightarrow{i} b, s)$ consisting of a morphism in $\A$ and an
object $s \in F(b)$;
\item a morphism from $(a \xrightarrow{i} b, s)$ to $(a' \xrightarrow{i'} b', s')$ is a triple $(f,g,h)$ where $f \maps a \to a'$ and $g \maps b \to b'$ are morphisms in $\A$ such that this square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
and $h \maps F(g)(s) \to s'$ is a morphism in $F(b')$.
\end{itemize}
As in any comma category we have a functor $\Phi \maps \A/U \to \A$ given on objects by $\Phi(a \xrightarrow{i} b, x)=a$ and on morphisms by $\Phi(f,g,h)=f$.

@@

The natural transformation $\theta$ gives a morphism
\[    \theta \maps   (c \xrightarrow{\psi} c+_a e, F(\psi)(s)) \; \to \; (c \xrightarrow{\phi f} d+_b e, F(\phi)(s))  \]
in $(\A/U)_c$.   Concretely, this arises from the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$e$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}e$};
\node (E) at (-1.5,0) {$a$};
\node (F) at (-1.5,-1) {$c$};
\node (G) at (2.5,-2) {$c+_{a}e$};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$k$} (F)
(E) edge node [above] {$h$} (A)
(F) edge node [below] {$f$} (C)
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\phi$} (D)
(F) edge [bend right] node[below]{$\psi$} (G)
(G) edge [dashed] node[below left]{$\exists ! \;\zeta $} (D)
(B) edge [bend left] (G);
\end{tikzpicture}
\]
Namely, we have $\theta = (1, \zeta, 1)$.   But when the original square in $\A$ (at upper left 
above) is a pushout, its pasting with the other pushout gives a pushout, so $\zeta$ and hence $\theta$ is isomorphism.

@@

(G) edge node [above] {$x$} (H)
(G) edge node [below] {$x'$} (I)

@@

(L) edge node[above]{$E(x)$} (B)
(L) edge node[below]{$E(x')$} (D)

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define $G \maps \A/U \to \A'/U'$ as follows.  On objects $G$ is given by
\[  G(a \xrightarrow{i} b, s) \; = \; (H(a) \xrightarrow{H(i)} H(b), \underline{s}), \]
where to define $\underline{x}$ we treat $s \in F(b)$ as a functor 
$x \maps 1 \to F(b)$ and let $\underline{s} \in F'(H(b))$ be the composite functor
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
On morphisms $(f,g,h) \maps (a \xrightarrow{i} b, s) \; \to \; 
(a' \xrightarrow{i'} b', s')$, $G$ is given by
\[   G(f,g,h) = (H(f),H(g),\underline{h}) \]
where to define $\underline{h}$ we treat $h \maps F(g)(s) \to s'$ as a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(b)$};
\node (I) at (-2,-1) {$F(b')$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$x$} (H)
(G) edge node [below] {$x'$} (I)
(H) edge node [right] {$F(g)$} (I);
\end{tikzpicture}
\]
and let $\underline{h} \maps F'(H(g))(\underline{s}) \to \underline{s}'$ be this
natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(b))$};
\node (D) at (2,-1) {$E(F(b'))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(b))$};
\node (F) at (4,-1) {$F^\prime(H(b'))$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi_0$} (L)
(B) edge node [above] {$\theta_b$} (E)
(D) edge node [above] {$\theta_{b'}$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(x)$} (B)
(L) edge node[below]{$E(x')$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]
One can check that with these definitions $G$ is a functor.

To make $G$ into a monoidal functor, we need to equip it with a laxator
\[  \gamma \maps G(a \xrightarrow{i} b, s) \otimes G(a' \xrightarrow{i'} b', s') \; 
\to \; G((a \xrightarrow{i} b, s) \otimes (a' \xrightarrow{i'} b', s')).\]
Note that
\[   G(a \xrightarrow{i} b, s) \otimes G(a' \xrightarrow{i'} b', s') = 
\left(H(a) + H(a') \xrightarrow{H(i) + H(i')} H(b) + H(b') , \; \phi'_{H(b), H(b')}(\underline{s},
\underline{s}')\right) \]
where $\phi'$ is the laxator for $F'$, while
\[    G((a \xrightarrow{i} b, s) \otimes (a' \xrightarrow{i'} b', s')) =
\left(H(a+a') \xrightarrow{H(i+i')} H(b+b'), \; \underline{\phi_{b,b'}(s,s')} \right)
\]
where $\phi$ is the laxator for $F$.  Since $H$ preserves finite colimits, the laxator
$\gamma$ is obvious except for the morphism from $\phi'_{H(b), H(b')}(\underline{s},
\underline{s}')$ to $\underline{\phi_{b,b'}(s,s')}$.   This is given by the following natural
transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node () at (3,-0.5) {$\scriptscriptstyle{\stackrel{E_{s,s'}} \cong}$};
\node () at (1.5,-0.5) {$\scriptscriptstyle{\stackrel{\lambda} \cong}$};
\node (L1) at (1,0) {$\scriptscriptstyle{1 \times E(1)}$};
\node (L2) at (1,-1) {$\scriptscriptstyle{E(1)}$};
\node (K) at (0,0) {$\scriptscriptstyle{1 \times 1}$};
\node (K2) at (0,-1) {$\scriptscriptstyle{1}$};
\node (B) at (2,0) {$\scriptscriptstyle{E(1) \times E(1)}$};
\node (D) at (2,-1) {$\scriptscriptstyle{E(1 \times 1)}$};
\node (E) at (4,0) {$\scriptscriptstyle{E(F(b)) \times E(F(b'))}$};
\node (F) at (4,-1) {$\scriptscriptstyle{E(F(b) \times F(b'))}$};
\node (G) at (6,0) {$\scriptscriptstyle{F'(H(b)) \times F'(H(b'))}$};
\node (H) at (8,0) {$\scriptscriptstyle{F'(H(b)+H(b'))}$};
\node (I) at (6,-1) {$\scriptscriptstyle{E(F(b+b'))}$};
\node (J) at (8,-1) {$\scriptscriptstyle{F'(H(b+b'))}$};
\path[->,font=\scriptsize,>=angle 90]
(K) edge node [left] {$\lambda$} (K2)
(K) edge node [above] {$1 \times E_0$} (L1)
(K2) edge node [below] {$E_0$} (L2)
(H) edge node [right] {$F'(H_{b,b'})$.} (J)
(E) edge node [above] {$\theta_b \times \theta_{b'}$} (G)
(G) edge node [above] {$\phi'_{H(b),H(b')}$} (H) 
(F) edge node [below] {$E(\phi_{b,b'})$} (I)
(I) edge node [below] {$\theta_{b+b'}$} (J)
(B) edge node [above] {$E(s) \times E(s')$} (E)
(D) edge node [below] {$E(s \times s')$} (F)
(E) edge node [right] {$E_{F(b),F(b')}$} (F)
(L1) edge node[above]{$E_0 \times 1$} (B)
(L1) edge node [left] {$\lambda$} (L2)
(L2) edge node[below]{$E(\lambda^{-1})$} (D)
(B) edge node[right]{$E_{1,1}$} (D);
\end{tikzpicture}
\]
We also need a natural transformation expressing lax preservation of the unit object by $G$, which
is defined similarly.   One can show that this laxator and unit morphism satisfy the coherence laws of a lax monoidal functor, and they are invertible, so $G$ is strong monoidal functor.   \john{If furthermore $F, F'$ and $E$ are symmetric, one can check that $G$ is symmetric monoidal, and that $\lH$ is a symmetric monoidal double functor.}
\end{proof}

@@

\begin{proof}
 This is \cite[Proposition~4.4]{Gray}.    Suppose each fiber $\X_a$ of the opfibration $U$ has an initial object $\bot_a$ and these objects are preserved (up to isomorphism) by the reindexing functors.   Define $L \maps \A \to \X$ on objects $a \in \A$ by $L(a) = \bot_a$.   Given a morphism $f\maps a\to a'$ in $\A$, define $L(f)$ to be the composite
 \begin{equation}\label{eq:Lonarrows}
  \bot_a\xrightarrow{\mathrm{Cocart}(f,\bot_a)}f_!(\bot_a)\xrightarrow{\;\chi_a\;}\bot_{a'}
 \end{equation}
where $\mathrm{Cocart}(f,\bot_a)$ is the cocartesian lifting of $f$ to $\bot_a$ and $\chi_a$ is the unique isomorphism between two initial objects in the fiber above $a'$.    The functor $L$ then becomes left adjoint to $R$ with unit $\iota_a \maps a \to U(L(a))$ being the identity, using the fact that $U(L(a)) = U(\bot_a) = a$.
\end{proof}

@@

Given the hypothesis of \cref{thm:equiv}, the isomorphism from $F\lCsp$ to ${}_L\lCsp(\inta F)$ works as follows.   It is the identity on objects and vertical morphisms, 
which are just objects and morphisms of $\A$.   It maps the decorated cospan
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$,};
\node (D) at (3,0) {$s \in F(m)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
to the structured cospan
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(a,I_a)$};
\node (B) at (1.5,0) {$(m,s)$};
\node (C) at (3,0) {$(b,I_b)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i,!)$} (B)
(C) edge node[above]{$(o,!)$} (B);
\end{tikzpicture}
\]
Here $I_a \in F(a)$ is the \define{trivial decoration} given by the composite
\[  \one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a), \]
while $! \maps F(i)(I_a) \to s$ is the unique morphism with this domain and
codomain (recall that $F(i)$ preserves initial objects), and similarly for
$! \maps F(o)(I_b) \to s$.   Finally, this isomorphism sends the map of decorated cospans
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\node (G) at (1,-1) {$\tau \maps F(h)(s) \to s'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
to this map of structured cospans:
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$(a,I_a)$};
\node (A') at (0,-0.5) {$(a',I_{a'}$};
\node (B) at (1.5,0.5) {$(m,s)$};
\node (C) at (3,0.5) {$(b,I_b)$};
\node (C') at (3,-0.5) {$(b',I_{b'}$.};
\node (D) at (1.5,-0.5) {$(m',s')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i,!)$} (B)
(C) edge node[above]{$(o,!)$} (B)
(A) edge node[left]{$L(f)$} (A')
(C) edge node[right]{$L(g)$} (C')
(A') edge node[below] {$(i',!)$} (D)
(C') edge node[below] {$(o',!)$} (D)
(B) edge node [left] {$(h,\tau)$} (D);
\end{tikzpicture}
\]

>>>>

In \cref{DecCospansDoubleCat} we construct the double category of decorated cospans, $F\lCsp$, and show how to construct maps between decorated cospan double categories. In \cref{Structured} we give a new construction of the double category of structured cospans, ${}_L \lCsp(\X)$.  In \cref{EquivDoubleCats} we prove that the double categories of decorated cospans and structured cospans are isomorphic under suitable conditions.  In \cref{spinoffs} we establish the isomorphism between structured and decorated cospans at the level of bicategories and categories (via decategorification).  In \cref{Applications} we describe applications, and show that open dynamical system can be described using decorated cospans but not structured cospans.

@@

We thank Daniel Cicala, Brendan Fong and Joe Moeller for helpful conversations.  We especially thank Michael Shulman for suggesting a better proof strategy that greatly improved this paper.  The third author would like to thank the General Secretariat for Research  and Technology (GSRT) and the Hellenic Foundation for Research and Innovation (HFRI).

@@

\begin{thm}\label{thm:decorated_cospans_1}
Let $\A$ be a category with finite colimits and $(F,\phi,\phi_0) \maps (\A,+) \to (\Cat,\times)$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $s \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a commutative diagram in $\A$ of the form
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a \define{decoration morphism} $\tau \maps F(h)(s) \to s'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

\begin{proof}[Proof of \cref{thm:decorated_cospans_1}] 
Let $U \maps \inta F \to \A$ be the monoidal opfibration associated to the lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ via \cref{lem:MonGroth}.  To obtain the desired double category we apply \cref{lem:shulman} to the functor $\Phi \maps \A/U \to \A$, which we now describe in more detail.  In the comma category $\A/U$:
\begin{itemize}
\item an object is a pair $(a \xrightarrow{i} b, s)$ consisting of a morphism in $\A$ and an
object $s \in F(b)$;
\item a morphism from $(a \xrightarrow{i} b, s)$ to $(a' \xrightarrow{i'} b', s')$ is a triple $(f,g,h)$ where $f \maps a \to a'$ and $g \maps b \to b'$ are morphisms in $\A$ such that this square commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1.5,0) {$a'$};
\node (C) at (0,-1) {$b$};
\node (D) at (1.5,-1) {$b'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(A) edge node[left]{$i$} (C)
(B) edge node[right]{$i'$} (D)
(C) edge node[below]{$g$} (D);
\end{tikzpicture}
\]
and $h \maps F(g)(s) \to s'$ is a morphism in $F(b')$.
\end{itemize}
As in any comma category we have a functor $\Phi \maps \A/U \to \A$ given on objects by $\Phi(a \xrightarrow{i} b, s)=a$ and on morphisms by $\Phi(f,g,h)=f$.

@@

The natural transformation $\theta$ gives a morphism
\[    \theta \maps   (c \xrightarrow{\psi} c+_a e, F(\psi)(s)) \; \to \; (c \xrightarrow{\phi f} d+_b e, F(\phi)(s))  \]
in $(\A/U)_c$.   Concretely, this arises from the universal property of the pushout:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$b$};
\node (B) at (1.5,0) {$e$};
\node (C) at (0,-1) {$d$};
\node (D) at (1.5,-1) {$d+_{b}e$};
\node (E) at (-1.5,0) {$a$};
\node (F) at (-1.5,-1) {$c$};
\node (G) at (2.5,-2) {$c+_{a}e$};
\node at (1.2,-0.8) {\Large{$\ulcorner$}};
\path[->,font=\scriptsize,>=angle 90]
(E) edge node [left] {$k$} (F)
(E) edge node [above] {$h$} (A)
(F) edge node [below] {$f$} (C)
(A) edge node[above]{$i$} (B)
(A) edge node[left]{$g$} (C)
(B) edge node[right]{$$} (D)
(C) edge node[below]{$\phi$} (D)
(F) edge [bend right] node[below]{$\psi$} (G)
(G) edge [dashed] node[below left]{$\exists ! \;\zeta $} (D)
(B) edge [bend left] (G);
\end{tikzpicture}
\]
Namely, we have $\theta = (1, \zeta, 1)$.   But when the original square in $\A$ (at upper left 
above) is a pushout, its pasting with the other pushout gives a pushout, so $\zeta$ and hence $\theta$ is an isomorphism.

@@

(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)

@@

(L) edge node[above]{$E(s)$} (B)
(L) edge node[below]{$E(s')$} (D)

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define $G \maps \A/U \to \A'/U'$ as follows.  On objects $G$ is given by
\[  G(a \xrightarrow{i} b, s) \; = \; (H(a) \xrightarrow{H(i)} H(b), \underline{s}), \]
where to define $\underline{x}$ we treat $s \in F(b)$ as a functor 
$s \maps 1 \to F(b)$ and let $\underline{s} \in F'(H(b))$ be the composite functor
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
On morphisms $(f,g,h) \maps (a \xrightarrow{i} b, s) \; \to \; 
(a' \xrightarrow{i'} b', s')$, $G$ is given by
\[   G(f,g,h) = (H(f),H(g),\underline{h}) \]
where to define $\underline{h}$ we treat $h \maps F(g)(s) \to s'$ as a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(b)$};
\node (I) at (-2,-1) {$F(b')$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)
(H) edge node [right] {$F(g)$} (I);
\end{tikzpicture}
\]
and let $\underline{h} \maps F'(H(g))(\underline{s}) \to \underline{s}'$ be this
natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(b))$};
\node (D) at (2,-1) {$E(F(b'))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(b))$};
\node (F) at (4,-1) {$F^\prime(H(b'))$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi_0$} (L)
(B) edge node [above] {$\theta_b$} (E)
(D) edge node [above] {$\theta_{b'}$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(s)$} (B)
(L) edge node[below]{$E(s')$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]
One can check that with these definitions $G$ is a functor.

To make $G$ into a monoidal functor, we need to equip it with a laxator
\[  \gamma \maps G(a \xrightarrow{i} b, s) \otimes G(a' \xrightarrow{i'} b', s') \; 
\to \; G((a \xrightarrow{i} b, s) \otimes (a' \xrightarrow{i'} b', s')).\]
Note that
\[   G(a \xrightarrow{i} b, s) \otimes G(a' \xrightarrow{i'} b', s') = 
\left(H(a) + H(a') \xrightarrow{H(i) + H(i')} H(b) + H(b') , \; \phi'_{H(b), H(b')}(\underline{s},
\underline{s}')\right) \]
where $\phi'$ is the laxator for $F'$, while
\[    G((a \xrightarrow{i} b, s) \otimes (a' \xrightarrow{i'} b', s')) =
\left(H(a+a') \xrightarrow{H(i+i')} H(b+b'), \; \underline{\phi_{b,b'}(s,s')} \right)
\]
where $\phi$ is the laxator for $F$.  Since $H$ preserves finite colimits, the laxator
$\gamma$ is obvious except for the morphism from $\phi'_{H(b), H(b')}(\underline{s},
\underline{s}')$ to $\underline{\phi_{b,b'}(s,s')}$.   This is given by the following natural
transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node () at (3,-0.5) {$\scriptscriptstyle{\stackrel{E_{s,s'}} \cong}$};
\node () at (1.5,-0.5) {$\scriptscriptstyle{\stackrel{\lambda} \cong}$};
\node (L1) at (1,0) {$\scriptscriptstyle{1 \times E(1)}$};
\node (L2) at (1,-1) {$\scriptscriptstyle{E(1)}$};
\node (K) at (0,0) {$\scriptscriptstyle{1 \times 1}$};
\node (K2) at (0,-1) {$\scriptscriptstyle{1}$};
\node (B) at (2,0) {$\scriptscriptstyle{E(1) \times E(1)}$};
\node (D) at (2,-1) {$\scriptscriptstyle{E(1 \times 1)}$};
\node (E) at (4,0) {$\scriptscriptstyle{E(F(b)) \times E(F(b'))}$};
\node (F) at (4,-1) {$\scriptscriptstyle{E(F(b) \times F(b'))}$};
\node (G) at (6,0) {$\scriptscriptstyle{F'(H(b)) \times F'(H(b'))}$};
\node (H) at (8,0) {$\scriptscriptstyle{F'(H(b)+H(b'))}$};
\node (I) at (6,-1) {$\scriptscriptstyle{E(F(b+b'))}$};
\node (J) at (8,-1) {$\scriptscriptstyle{F'(H(b+b'))}$};
\path[->,font=\scriptsize,>=angle 90]
(K) edge node [left] {$\lambda$} (K2)
(K) edge node [above] {$1 \times E_0$} (L1)
(K2) edge node [below] {$E_0$} (L2)
(H) edge node [right] {$F'(H_{b,b'})$.} (J)
(E) edge node [above] {$\theta_b \times \theta_{b'}$} (G)
(G) edge node [above] {$\phi'_{H(b),H(b')}$} (H) 
(F) edge node [below] {$E(\phi_{b,b'})$} (I)
(I) edge node [below] {$\theta_{b+b'}$} (J)
(B) edge node [above] {$E(s) \times E(s')$} (E)
(D) edge node [below] {$E(s \times s')$} (F)
(E) edge node [right] {$E_{F(b),F(b')}$} (F)
(L1) edge node[above]{$E_0 \times 1$} (B)
(L1) edge node [left] {$\lambda$} (L2)
(L2) edge node[below]{$E(\lambda^{-1})$} (D)
(B) edge node[right]{$E_{1,1}$} (D);
\end{tikzpicture}
\]
We also need a natural transformation expressing lax preservation of the unit object by $G$, which
is defined similarly.   One can show that this laxator and unit morphism satisfy the coherence laws of a lax monoidal functor, and they are invertible, so $G$ is strong monoidal functor.   If furthermore $F, F'$ and $E$ are symmetric, one can check that $G$ is symmetric monoidal, and that $\lH$ is a symmetric monoidal double functor.
\end{proof}

@@

\begin{proof}
 This is \cite[Proposition~4.4]{Gray}.    Suppose each fiber $\X_a$ of the opfibration $U$ has an initial object $\bot_a$ and these objects are preserved (up to isomorphism) by the reindexing functors.   Define $L \maps \A \to \X$ on objects $a \in \A$ by $L(a) = \bot_a$.   Given a morphism $f\maps a\to a'$ in $\A$, define $L(f)$ to be the composite
\[
  \bot_a\xrightarrow{\mathrm{Cocart}(f,\bot_a)}f_!(\bot_a)\xrightarrow{\;\chi_a\;}\bot_{a'}
\]
where $\mathrm{Cocart}(f,\bot_a)$ is the cocartesian lifting of $f$ to $\bot_a$ and $\chi_a$ is the unique isomorphism between two initial objects in the fiber above $a'$.    The functor $L$ then becomes left adjoint to $R$ with unit $\iota_a \maps a \to U(L(a))$ being the identity, using the fact that $U(L(a)) = U(\bot_a) = a$.
\end{proof}

@@

Given the hypothesis of \cref{thm:equiv}, the isomorphism from $F\lCsp$ to ${}_L\lCsp(\inta F)$ works as follows.   It is the identity on objects and vertical morphisms, 
which are just objects and morphisms of $\A$.   It maps the decorated cospan
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$,};
\node (D) at (3,0) {$s \in F(m)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
to the structured cospan
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(a,I_a)$};
\node (B) at (1.5,0) {$(m,s)$};
\node (C) at (3,0) {$(b,I_b)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i,!)$} (B)
(C) edge node[above]{$(o,!)$} (B);
\end{tikzpicture}
\]
Here $I_a \in F(a)$ is the \define{trivial decoration} given by the composite
\[  \one\xrightarrow{\phi_0}F(0)\xrightarrow{F(!_a)}F(a), \]
while $! \maps F(i)(I_a) \to s$ is the unique morphism with this domain and
codomain (recall that $F(i)$ preserves initial objects), and similarly for
$! \maps F(o)(I_b) \to s$.   Finally, this isomorphism sends the map of decorated cospans:
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\node (G) at (1,-1) {$\tau \maps F(h)(s) \to s'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[below] {$i'$} (D)
(C') edge node[below] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
to this map of structured cospans:
\[
\begin{tikzpicture}[scale=1.5,baseline=(current bounding box.center)]
\node (A) at (0,0.5) {$(a,I_a)$};
\node (A') at (0,-0.5) {$(a',I_{a'})$};
\node (B) at (1.5,0.5) {$(m,s)$};
\node (C) at (3,0.5) {$(b,I_b)$};
\node (C') at (3,-0.5) {$(b',I_{b'})$.};
\node (D) at (1.5,-0.5) {$(m',s')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i,!)$} (B)
(C) edge node[above]{$(o,!)$} (B)
(A) edge node[left]{$L(f)$} (A')
(C) edge node[right]{$L(g)$} (C')
(A') edge node[below] {$(i',!)$} (D)
(C') edge node[below] {$(o',!)$} (D)
(B) edge node [left] {$(h,\tau)$} (D);
\end{tikzpicture}
\]
